---
title: "ADSS Assessment: C. Clustering"
author: "Kevin Donkers (EI CDT) - 700063874"
date: "16/01/2021"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: false
    toc: false
  bookdown::html_document2:
    fig_caption: yes
    number_sections: false
    smart: no
---
```{r setup, include=FALSE}
library(here)
library(bookdown)
knitr::opts_chunk$set(echo=FALSE, message=FALSE, cache=TRUE, warning=FALSE, include=FALSE)
# knitr::opts_chunk$set(echo=TRUE, message=TRUE, cache=FALSE, warning=FALSE, include=TRUE)
knitr::read_chunk(here::here('scripts/init.R'))
knitr::read_chunk(here::here('scripts/question1.R'))
knitr::read_chunk(here::here('scripts/question2.R'))
knitr::read_chunk(here::here('scripts/question3.R'))
```

<!-- Fig link: \@ref(fig:TagNoSpecialChars) -->
# 0. Setup

Code for wrangling data:

```{r 0LoadData, include=TRUE, echo=TRUE}
```

# Q1. Initial data analysis tasks.

```{r 1Init}
```

## (i) Summarise the data in the Characteristics.csv dataset, and plot the distributions for the percentage of industrial and commercial customers, transformer ratings and pole or ground monitored substations.

I was plotted a pairs plot using the `GGally` function `ggpairs()`. This allowed comparison of percentage industrial/commercial customers vs transformer ratings, grouped by pole/ground label (see Figure \@ref(fig:1iGgpairs)).

```{r 1iGgpairs, include=TRUE, echo=TRUE, fig.cap="Summary of percentage industrial/commercial customers vs transformer ratings, grouped by pole/group transformer labels."}
```

Here is what I found from the pairs plot:
- Pole type transformers tend to be of lower rating to groud transformers
- There is not much of a correlation between the percentage of non-domestic use to the type of transformer.
- Non-domestic use is more likely with ground than pole type tranformers



## (ii) Using this and other analyses you think appropriate, describe the relationships between the different substation characteristics (transformer type, number of customers, rating, percentage of I&C customers and number of feeders).

I plotted anothers pairs plot which allowed further comparison with number of customers, rating and number of feeders (see Figure \@ref(fig:1iiGgpairs)).

```{r 1iiGgpairs, include=TRUE, echo=TRUE, fig.cap="Summary of substation characteristics, grouped by pole/group transformer labels."}
```

Here are my comments:
- Pole transformers tend to have fewer customers, fewer feeders and a lower rating compared to ground transformers
- The number of feeders correlates positively with the number of customers
- The fewer the number of customers the more likely they are non-domestic. Less strong for pole transformers
- Higher rated transformers tend to have more LV-feeders


\pagebreak
# Q2. Initial clustering tasks
## Using the scaled daily measurements from the January_2013 dataset perform hierarchical clustering for the daily average demand (that is each 10 minute period for the given substation should be averaged over all days where measurement is available):

Code for scaling daily averages:
```{r 2Init, include=TRUE, echo=TRUE}
```

## (i) Using your preferred choice of a dissimilarity function, create a distance matrix for these data and produce a dendrogram.

Here I used the `dist()` function to determine the disimilarity of the substations.

```{r 2iDist, include=TRUE, echo=TRUE}
```

I could then plot a dendogram (Figure \@ref(fig:2iDendPlot)):

```{r 2iDendPlot, include=TRUE, echo=TRUE, fig.cap="Dendogram of substation data"}
```


## (ii) Choose an appropriate number of clusters and label each substation according to its cluster membership.

We can cut the dendogram using `cutree()` to the number of clusters we want (in this case I chose 7). Then we can assign these cluster values to their corresponding substation.

```{r 2iiClusters, include=TRUE, echo=TRUE}
```

## (iii) For each of your clusters, plot the daily average demand for 1) All days, 2) Weekdays, 3) Saturdays and 4) Sundays.

Due to time constraints, I was only able to plot the daily average demand for "1) All days", which is in Figure \@ref(fig:2iiiAlldays):

```{r 2iiiAlldays, include=TRUE, fig.cap="The seven clusters of usage profiles of all the substations, for all days of the week."}
```

We can see similarities in shape between some clusters (e.g. 2, 4 and 5 have similar profile shapes but differing average demand). We can see that the clustering has managed to pick up unusual profiles with very few members, such as cluster 7 which has a peak demand some time around 1-3am. 


## (iv) Produce summaries of the variables in Characteristics.csv for each of your clusters.

Did not answer - ran out of time

## (v) Desribe your clusters based on the information in Characteristics.csv and choose names for them. Describe the patterns of their power demands for each cluster.

Did not answer - ran out of time




\pagebreak
# Q3. Allocating new substations
## The Dataset NewSubstations.csv contains information for five new substations.


## (i) For each substation, on the same plot, plot the daily average demand for 1) All days, 2) Weekdays, 3) Saturdays and 4) Sundays (one plot per new substation).

Did not answer - ran out of time

## (ii) Using k-means (or other version, i.e. based on medians), allocate these new substations to one of your clusters

I quickly ran a `kmeans()` clustering on the new data but because there are only 5 entries I was not able to cluster to the 7 clusters identified in hierarchical clustering.

Therefore I figured it would be best ot merge the old and new substation datasets and run `kmeans()` on this merged dataset, aiming to identify 7 clusters and then finding which clusters the five entries from the new dataset ended up in.

However, I ran out of time to wrangle the new data to be mergeable with the old data, given they have 144 columns with slightly differing names.

I also thought it might be possible to pull out some characteristics of the hierarchical clustering (e.g. cluster centres) and used that with a `kmeans()` call on the five new data points, but I couldn't work out how to do that.

## (iii) Based on your summaries and plots, is the cluster allocation as you expected?

Did not answer - ran out of time

\pagebreak
# Q4. Exploring differences between years

My apologies, I ran out of time to complete this question.