---
title: "ADSS Assessment: C. Clustering"
author: "Kevin Donkers (EI CDT) - 700063874"
date: "16/01/2021"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    number_sections: false
    toc: false
  bookdown::html_document2:
    fig_caption: yes
    number_sections: false
    smart: no
---
```{r setup, include=FALSE}
library(here)
library(bookdown)
knitr::opts_chunk$set(echo=FALSE, message=FALSE, cache=TRUE, warning=FALSE, include=FALSE)
# knitr::opts_chunk$set(echo=TRUE, message=TRUE, cache=FALSE, warning=FALSE, include=TRUE)
knitr::read_chunk(here::here('scripts/init.R'))
knitr::read_chunk(here::here('scripts/question1.R'))
knitr::read_chunk(here::here('scripts/question2.R'))
knitr::read_chunk(here::here('scripts/question3.R'))
```

<!-- Fig link: \@ref(fig:1.summary) -->

```{r 0.load_data}
```

# Q1. Initial data analysis tasks.

```{r 1.init}
```

## (i) Summarise the data in the Characteristics.csv dataset, and plot the distributions for the percentage of industrial and commercial customers, transformer ratings and pole or ground monitored substations.


```{r 1.i.ggpairs}
```

- Pole type transformers tend to be of lower rating to groud transformers
- There is not much of a correlation between the percentage of non-domestic use to the type of transformer.
- Non-domestic use is more likely with ground than pole type tranformers


```{r 1.ii.ggpairs}
```


- Pole transformers tent to have fewer customers, fewer feeders and a lower rating compared to ground transformers
- The number of feeders correlates positively with the number of customers
- The fewer the number of customers the more likely they are non-domestic. Less strong for pole transformers
- Higher rated transformers tend to have more LV-feeders
- 


## (ii) Using this and other analyses you think appropriate, describe the relationships between the different substation characteristics (transformer type, number of customers, rating, percentage of I&C customers and number of feeders).



\pagebreak
# Q2. Initial clustering tasks
## Using the scaled daily measurements from the January_2013 dataset perform hierarchical clustering for the daily average demand (that is each 10 minute period for the given substation should be averaged over all days where measurement is available):

```{r 2.init}
```

## (i) Using your preferred choice of a dissimilarity function, create a distance matrix for these data and produce a dendrogram.


## (ii) Choose an appropriate number of clusters and label each substation according to its cluster membership.


## (iii) For each of your clusters, plot the daily average demand for 1) All days, 2) Weekdays, 3) Saturdays and 4) Sundays.


## (iv) Produce summaries of the variables in Characteristics.csv for each of your clusters.



## (v) Desribe your clusters based on the information in Characteristics.csv and choose names for them. Describe the patterns of their power demands for each cluster.






\pagebreak
# Q3. Allocating new substations
## The Dataset NewSubstations.csv contains information for five new substations.

```{r 3.init}
```
## (i) For each substation, on the same plot, plot the daily average demand for 1) All days, 2) Weekdays, 3) Saturdays and 4) Sundays (one plot per new substation).


## (ii) Using k-means (or other version, i.e. based on medians), allocate these new substations to one of your clusters


## (iii) Based on your summaries and plots, is the cluster allocation as you expected?

\pagebreak
# Q4. Exploring differences between years

My apologies, I ran out of time to complete this question.